US President Prediction

### 2024-10-30 paper/paper.qmd
Updated model and result

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
```

@@

# Model {#set-model}

As indicated in the @sec-data, we do not have a confident belief of priors of parameters of our model since Harris replace the role of Biden by indecent and we do not have enough data about her previous polling outcomes as a president candidate compared to Biden. Therefore, we would not use the bayesian model but instead using a more simple multilinear regression model which predicts the support rate of Trump and Harris using time towards elections date and pollsters as they are beliveved to highly corelated to the response variable in @sec-data. We would bulid 8 models whihc represent national trend and that of seven swing states. 
## Model set-up

Define $y_{i,\text{stateA}}$ as the percentage of support for a candicate in a poll in the specific state.

$$
y_{i,national} = \beta_{0,national} + \beta_{1,time,national} + \beta_{2,pollster,national} + \epsilon_{national}
$$ 
$$
y_{i,\text{AZ}} = \beta_{0,\text{AZ}} + \beta_{1,\text{time,AZ}} + \beta_{2,\text{pollster,AZ}} + \epsilon_{\text{AZ}}
$$
$$
y_{i,\text{FL}} = \beta_{0,\text{FL}} + \beta_{1,\text{time,FL}} + \beta_{2,\text{pollster,FL}} + \epsilon_{\text{FL}}
$$
$$
y_{i,\text{GA}} = \beta_{0,\text{GA}} + \beta_{1,\text{time,GA}} + \beta_{2,\text{pollster,GA}} + \epsilon_{\text{GA}}
$$
$$
y_{i,\text{MI}} = \beta_{0,\text{MI}} + \beta_{1,\text{time,MI}} + \beta_{2,\text{pollster,MI}} + \epsilon_{\text{MI}}
$$
$$
y_{i,\text{NC}} = \beta_{0,\text{NC}} + \beta_{1,\text{time,NC}} + \beta_{2,\text{pollster,NC}} + \epsilon_{\text{NC}}
$$
$$
y_{i,\text{PA}} = \beta_{0,\text{PA}} + \beta_{1,\text{time,PA}} + \beta_{2,\text{pollster,PA}} + \epsilon_{\text{PA}}
$$
$$
y_{i,\text{WI}} = \beta_{0,\text{WI}} + \beta_{1,\text{time,WI}} + \beta_{2,\text{pollster,WI}} + \epsilon_{\text{WI}}
$$

time:The date when the pollster starts polling

pollster: The pollster which run the polling

We run the model in R [@citeR].

### Model justification

# Results {#set-result}

@@

## pollster methodology analysis

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(ggpubr)
```

@@

# Model {#sec-model}

In this analysis, we aim to model the popularity trends for Kamala Harris and Donald Trump, based on high-quality polling data collected at the national level after Harris’s declaration on July 21, 2024. We used Bayesian linear regression models with Gaussian error structures to estimate changes in polling percentages over time for each candidate.

## Data Filtering and Preparation

The data were first filtered to retain only high-quality polls (with a numeric grade of 2.0 or above and a transparency score of 4 or above). We focused on polls where the `state` was either explicitly labeled as "National" or where the state information was missing (which we imputed as "National"). Additionally, we filtered polls collected after July 21, 2024, when Kamala Harris announced her candidacy.

- **Kamala Harris Data**: After filtering for high-quality polls and setting the timeframe, we calculated the number of Harris supporters per poll by taking the product of her polling percentage and the poll’s sample size.
- **Donald Trump Data**: We applied the same filtering and calculation methods to Trump’s data, ensuring comparability between both candidates.

To avoid issues with missing values in our models, we excluded any rows where key variables (e.g., polling percentage or end date) were missing.

## Model Specification

Two separate Bayesian linear regression models were fitted using the `stan_glm` function. Both models specified the formula:

\[
\text{pct} \sim \text{end\_date}
\]

indicating that we are modeling the polling percentage as a function of time (end date of the polls).

### Model for Kamala Harris

The model for Kamala Harris was fitted as follows:

- **Response Variable**: `pct` (polling percentage)
- **Predictor**: `end_date` (date of the poll)
- **Family**: Gaussian (normal distribution)
- **Priors**:
  - Normal prior for the slope: \( \text{Normal}(0, 0.1) \)
  - Normal prior for the intercept: \( \text{Normal}(50, 5) \)
  - Exponential prior for auxiliary parameters: \( \text{Exponential}(1) \)

### Model for Donald Trump
The model for Donald Trump was fitted using the same specifications as for Kamala Harris

# Results {#set-result}
The results of the Bayesian linear regression models for Kamala Harris and Donald Trump are summarized in @tbl-model_summary_H and @tbl-model_summary_T.

```{r}
#| label: tbl-model_summary_H
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables
library(modelsummary)

# Load the saved models
model1 <- readRDS(file = here::here("models/models_harris.rds"))

# Extract a tidy summary of the model
model_summary <- broom.mixed::tidy(model1, effects = "fixed", conf.int = TRUE) %>%
  select(term, estimate, std.error, conf.low, conf.high) %>%
  rename(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    `2.5% CI` = conf.low,
    `97.5% CI` = conf.high
  )

# Display the summary table
kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)

```
```{r}
#| label: tbl-model_summary_T
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables

# Load the saved models
model2 <- readRDS(file = here::here("models/models_trump.rds"))

# Extract a tidy summary of the model
model_summary <- broom.mixed::tidy(model2, effects = "fixed", conf.int = TRUE) %>%
  select(term, estimate, std.error, conf.low, conf.high) %>%
  rename(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    `2.5% CI` = conf.low,
    `97.5% CI` = conf.high
  )

# Display the summary table
kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)
```
## Overall trend
The intercepts for both models are significantly negative, indicating a downward trend in polling percentages over time for both candidates. The positive coefficients for end_date suggest a slight increase in polling percentages as the election date approaches. However, the magnitude of these increases is relatively small.

## Model Diagnostics
The posterior predictive checks (Figures 1 and 2) and the credible intervals for predictors (Figures 3 and 4) indicate that the models fit the data well. The diagnostics for both Harris and Trump models show that the models are stable and the predictions are reliable.Further diagnostics (Figures 5 to 8) confirm the robustness of the models. The R-hat values are all below 1.1, indicating good convergence of the MCMC chains.

@@

## Model check

```{r}
#| label: fig-pos_check_H
#| echo: false
#| fig-cap: "Posterior Predictive Check for All Models"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

pp_check(model1) +  
    theme_classic() +
    theme(legend.position = "bottom")

```
```{r}
#| label: fig-pos_check_T
#| echo: false
#| fig-cap: "Posterior Predictive Check for All Models"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

pp_check(model2) +  
    theme_classic() +
    theme(legend.position = "bottom")

```
```{r}
#| label: fig-CI_H
#| echo: false
#| fig-cap: "CI for predictors for Harris"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

plot(
  model1,
  "areas")
```
```{r}
#| label: fig-CI_T
#| echo: false
#| fig-cap: "CI for predictors for Trump"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

plot(
  model2,
  "areas")
```

```{r}
#| label: fig-DI_H
#| echo: false
#| fig-cap: "Model diagonostic for Harris"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

plot(model1, "trace")
plot(model1, "rhat")

```
```{r}
#| label: fig-DI_T
#| echo: false
#| fig-cap: "Model diagonostic for Trump"

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots

plot(model2, "trace")
plot(model2, "rhat")

```
## pollster methodology analysis
